---
title: "Problem Set 1"
author: "73-265 Economics & Data Science"
date: "Fall 2022"
output: pdf_document
---

```{r}
## load data
boston <- read.csv("data/boston.csv", head = TRUE)
```
 
## Question 1

## Part 1-A
```{r}
## create two subsets
boston.treat <- subset(boston, treatment == 1)
boston.ctrl <- subset(boston, treatment == 0)
## quartiles
summary(boston.treat$income)
summary(boston.ctrl$income)  
# mean and standard deviation
c(mean(boston.treat$income),mean(boston.ctrl$income))
c(sd(boston.treat$income),sd(boston.ctrl$income))
# mean difference?
(mean(boston.treat$income)-mean(boston.ctrl$income))/mean(boston$income)
(mean(boston.treat$income)-mean(boston.ctrl$income))/sd(boston$income)
```
    
The distributions of household income across the control and treatment group are
balanced in the sense that the median is the same. However, one may argue that 
the difference in means is economically
significant because it is 11% of the average income in the sample
and 21% of the standard deviation.

## Part 1-B
```{r}
mean(boston.treat$male)
mean(boston.ctrl$male)
```

Of all the individuals assigned to the treatment group,
`r round(mean(boston.treat$male),3)*100`% are males. Of all the
individuals assigned to the control group,
`r round(mean(boston.ctrl$male),3)*100`%
are male. They differ by about
`r round((mean(boston.ctrl$male) -
  mean(boston.treat$male))*100,0)`
percentage points. The treatment and control groups are relatively
balanced on income and gender.  

## Question 2

```{r}
## compute change in attitude
boston$change <- boston$numberim.post - boston$numberim.pre
treat.change <- mean(boston$change[boston$treatment == 1],
                     na.rm = TRUE)
ctrl.change <- mean(boston$change[boston$treatment == 0],
                    na.rm = TRUE)
treat.change
ctrl.change
## average treatment effect
treat.change - ctrl.change
```
    
The average change within the control group is a decrease in
`r round(ctrl.change*-1, 2)` points on the 5 point scale (where a decrease
indicates more inclusionary attitudes). The average change within the
treatment group is an increase in `r round(treat.change, 2)` points on the 5
point scale (where an increase indicates more exclusionary
attitudes). The average treatment effect is the difference of the
two. On average, the responses of individuals in the treatment group
were more exclusionary than the control group by
`r round(treat.change - ctrl.change,2)` points. Exposure to
simulated demographic changes caused this increase in exclusionary
attitudes.

## Question 3

```{r}
t.col.change <- mean(boston$change[boston$treatment == 1 &
                                       boston$college == 1],
                     na.rm = TRUE)
c.col.change <- mean(boston$change[boston$treatment == 0 &
                                       boston$college == 1],
                     na.rm = TRUE)
t.nocol.change <- mean(boston$change[boston$treatment == 1 &
                                         boston$college == 0],
                       na.rm = TRUE)
c.nocol.change <- mean(boston$change[boston$treatment == 0 &
                                         boston$college == 0],
                       na.rm = TRUE)
ate.col <- t.col.change - c.col.change
ate.nocol <- t.nocol.change - c.nocol.change
ate.col
ate.nocol
ate.col - ate.nocol
```
    
The average treatment effect (using the 'change' variable) among
those with a college education is an increase in exclusionary
attitudes of about `r round(ate.col,2)` points. Among those
without a college education, there is an increase in exclusionary
attitudes of about `r round(ate.nocol,2)`. Both of these effects
are on a 5 point scale.

The difference between them is less than `r round(ate.col -
  ate.nocol,2)` points on this scale and is negligible. Because the
average treatment effects do not vary across the two subgroups, there
is no evidence of a differential impact.

## Question 4

## Part 4-A

```{r}
## calculate a vector of age quartiles
qrt <- quantile(boston$age, probs = seq(from = 0, to = 1, by = 0.25)) 
qrt
## create an age quartile group variable
boston$age.qrt<-1
boston$age.qrt[boston$age>qrt[2]&boston$age<=qrt[3]]<-2
boston$age.qrt[boston$age>qrt[3]&boston$age<=qrt[4]]<-3
boston$age.qrt[boston$age>qrt[4]]<-4
## treatment group change by age
t.age.change <- tapply(boston$change[boston$treatment == 1], 
                       boston$age.qrt[boston$treatment == 1], mean, 
                       na.rm = TRUE)
## control group change by age
c.age.change <- tapply(boston$change[boston$treatment == 0], 
                       boston$age.qrt[boston$treatment == 0], mean, 
                       na.rm = TRUE)
ate.age <- t.age.change - c.age.change
ate.age
```

## Part 4-B
```{r}
## treatment group change by ideology
t.idea.change <- tapply(boston$change[boston$treatment == 1],
                        boston$ideology[boston$treatment == 1],
                        mean, na.rm = TRUE)
## control group change by ideology
c.idea.change <- tapply(boston$change[boston$treatment == 0],
                        boston$ideology[boston$treatment == 0],
                        mean, na.rm = TRUE)
ate.idea <- t.idea.change - c.idea.change
ate.idea
```
    
Within each age category and ideology category, we see the largest
changes between treatment and control groups at the extremes of the
distribution. We also see a fair amount of difference in treatment
effects, suggesting there may be a differential impact by age and by
ideology.

    

